---
title: "Task 4: Naples Reef Purple Urchins (SILVA)"
author: "German Silva"
date: "10/28/2021"
output: 
  html_document: 
    number_sections: yes
---

# Overview

This code is for practicing exploratory data visualization with Purple Urchin data from Naples Reef in 2010. The data shows two different plots useful for exploratory data visualization. Data visualization is then followed with a descriptive stats summary table that can be used to discuss characteristics of the data.

**Data Source**: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

# Data Exploration

## Creation of a stored object for purple urchins found at Naples Reef in 2010

- Read in `urchins.csv` using `read_csv()` and `here()`
- Clean names with `clean_names()`
- Convert `date` column into `class = Date`
- Make `year` column with `year()`
- Subset for purple urchins at Naples reef in 2010 using `filter()`

```{r}
naples_reef_2010 <- read_csv(here("data", "urchins.csv")) %>% # read in the csv
  clean_names() %>% # converts column names unto lower_snakecase
  mutate(date = mdy(date)) %>%  # converts date column into a date class
  mutate(year = year(date)) %>% # add a year column
  filter(common_name == "Purple Urchin",
         year == "2010",
         site == "NAPL") # filtering to subset for the conditions we are looking for
```

## Exploratory Data Visualization

```{r}
ggplot(naples_reef_2010, aes(x= size)) + # opens data for graphing
  geom_histogram(bins = 10)+ # sets it as a histogram with 10 bins
  facet_wrap(~treatment) # facets it by treatment

ggplot(naples_reef_2010, aes(sample = size)) + # opens data
  geom_qq()+ # creates a q-q plot
  facet_wrap(~treatment) # facets the qq plot by treatment
```


**Takeaway Points**: 
Looking at both the histograms and the Q-Q plots, it appears that the data is fairly close to normally distributed (Q-Q plots have an overall linear trend). Urchins in the annual kelp removal treatment have a central tendency lower than that of the control group. 

## Summary Statistics Table

```{r}
naples_2010_stats <- naples_reef_2010 %>% 
  group_by(treatment) %>% 
  summarize(mean_diameter_cm = round(mean(size), 2),
            median_diameter_cm = round(median(size), 2),
            sd_cm = round(sd(size), 2),
            sample_size = n())

kable(naples_2010_stats, caption = "**Table 1** Central Tendency and Data Spread Statistics for  Purple Urchins at Naples Reef in 2010 by Treatment")
```

Urchins in the annual kelp removal treatment have on lower mean ($\bar{x}$ = `r naples_2010_stats$mean_diameter_cm[1]`) and median body diameter (mdn = `r naples_2010_stats$median_diameter_cm[1]`) than those in the control group ($\bar{x}$ = `r naples_2010_stats$mean_diameter_cm[2]`, mdn = `r naples_2010_stats$median_diameter_cm[2]`). Data for the annual kelp removal treatment also have a larger data spread (sd = `r naples_2010_stats$sd_cm[1]`) compared to the control group (sd = `r naples_2010_stats$sd_cm[2]`). 
