---
title: "Task 4: Naples Reef Purple Urchins (SILVA)"
author: "German Silva"
date: "10/28/2021"
output: 
  html_document: 
    number_sections: yes
---

# Overview



**Data Source**: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

# Data Exploration

## Create a stored object for purple urchins found at Naples Reef in 2010

- Read in `urchins.csv` using `read_csv()` and `here()`
- Clean names with `clean_names()`
- Convert `date` column into `class = Date`
- Make `year` column with `year()`
- Subset for purple urchins at Naples reef in 2010 using `filter()`

```{r}
naples_reef_2010 <- read_csv(here("data", "urchins.csv")) %>% # read in the csv
  clean_names() %>% # converts column names unto lower_snakecase
  mutate(date = mdy(date)) %>%  # converts date column into a date class
  mutate(year = year(date)) %>% # 
  filter(common_name == "Purple Urchin",
         year == "2010",
         site == "NAPL")
```

