---
title: 'Task 3: Santa Barbara Urchins (SILVA)'
author: "German Silva"
date: "10/23/2021"
output: html_document
---
# Overview


**Data Source**: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
```

### Data visualization of size distribution between two colors of urchin
- Read in `urchin.csv`
- Clean names
- Separate names to obtain color column 
- Group by color 
- Create histogram of size distribution
  - reduce the number of bins
```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% # read in the data
  clean_names() %>% # clean names into lower_snakecase with clean_names()
  separate(col = common_name, # state which column we want split up
           into= c("color_type", "animal"), # state what you want the new columns to be
           sep = " ", # which separator is being used to make the columns
           remove = FALSE) %>%  # we are not removing the original column
  group_by(color_type) # create groups based on color

ggplot(urchins, aes(x = size, fill = color_type))+
  geom_histogram(bins = 10, alpha= 0.5, position = "identity") +
  scale_color_manual(values = c("darkorchid4", "darkred")) +
  scale_fill_manual(values = c("darkorchid4", "darkred")) +
  labs(x = "Urchin Diameter in cm",
       y = "Number of Observations",
       fill = "Color of Urchin") +
  ggtitle("Difference in Size Distribution of Red and Purple Urchins (SILVA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

